---
title: "Lab2_Final_2"
author: "Luke Profio"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

# Setup

## Load libraries

```{r message=FALSE, warning=FALSE}
library(ranger)
library(tidyverse)
library(tidymodels)
library(DataExplorer)
library(forcats)
library(conflicted)
library(vip)
conflicted::conflicts_prefer(yardstick::spec)
```

## Read CSV

```{r p2 load_data}
df <- read_csv("scholastic_travel.csv", show_col_types = FALSE) |>
  mutate(across(where(is.character), as.factor))
```


## Set global variables to be used throughout the analysis


```{r p3}
custom_metrics <- metric_set(accuracy, sens, spec, precision, recall)
set.seed(42)
```


# Data Exploration

## Missing values

```{r p4}
plot_intro(df)
```
## Retention by Categorical Variables

```{r p5}
plot_bar(df, by = "RetainedLabel")
```

## Retention by Categorical Variables (Zoomed In)

```{r p6}
df_proportions <- df |>
  group_by(SPRNewExisting, RetainedLabel) |>
  summarise(Count = n()) |> 
  mutate(Proportion = Count / sum(Count)) |>
  ungroup() 

ggplot(df_proportions, aes(x = Proportion, y = SPRNewExisting, fill = RetainedLabel)) +
  geom_bar(stat = "identity", position = "fill", orientation = "y") +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "Proportion", y = "Program Code", fill = "Status") +
  ggtitle("Proportion of Churned vs Retained by SPRNewExisting") +
  theme_minimal()
```

```{r p7}
df_proportions
```

## Retention by Numeric Variables 

```{r p8}
ggplot(df, aes(x = RetainedLabel, y = FPP, fill = RetainedLabel)) +
  geom_boxplot() +
  labs(title = "Side by Side Boxplots", x = "Customer Retained Label", y = "Number of Full Paying Participants (FPP)") +
  theme_minimal()
```

```{r p9}
df |>
  group_by(RetainedLabel) |>
  summarise(count=n(), across(where(is.numeric), mean))
```

# Data Modeling

## Split Data for Training and Testing

```{r p10}
split <- initial_split(df, 
                       prop = 0.80, 
                       strata = RetainedLabel) 

train <- training(split)
test <- testing(split)
```

## K-Nearest Neighbor

```{r p11}
library(kknn)
library(caret)

knn_recipe <- 
  recipe(RetainedLabel ~ FromGrade + IsNonAnnual + Tuition + ToGrade + TotalSchoolEnrollment + FPP + SchoolSponsor + TotalPax + SingleGradeTripFlag, data = train) |>
  step_normalize(all_numeric_predictors()) |>
  step_dummy(all_nominal(), -all_outcomes())

knn_model <- 
  nearest_neighbor(weight_func = "cos", neighbors = 20, dist_power = 2) |>
  set_engine("kknn") |>
  set_mode("classification")

knn_workflow <- 
  workflow() |> 
  add_model(knn_model) |> 
  add_recipe(knn_recipe) |>
  fit(data = train)
```


## Evaluate the Model

```{r p12}
pred_class <- predict(knn_workflow,
                      new_data = test,
                      type = "class")

pred_probability <- predict(knn_workflow,
                      new_data = test,
                      type = "prob")

knn_results <- test |>
  bind_cols(pred_class, pred_probability)

custom_metrics(knn_results,
               truth = RetainedLabel,
               estimate = .pred_class)
```

### Confusion Matrix

```{r p13}
conf_mat(knn_results, truth = RetainedLabel,
         estimate = .pred_class)
```

### ROC Area Under the Curve

```{r p14}
roc_auc(knn_results,
        truth = RetainedLabel,
        .pred_Retained,
        event_level = "second")
```

### ROC Curve Visual

```{r p15}
knn_results |>
  roc_curve(truth = RetainedLabel, .pred_Retained, event_level = "second") |>
  autoplot()
```

## Random Forest

```{r p16}
cv_set <- vfold_cv(train, strata = RetainedLabel, v = 5)

rf_recipe <- 
  recipe(RetainedLabel ~ SPRNewExisting + FromGrade + InitialSystemDate + FRPActive + LatestRPL + FromGrade + FPPtoSchoolenrollment + SPRGroupRevenue + FRPTakeuppercent + DepartureDate + FPPtoPAX + DifferenceTraveltoFirstMeeting + DepositDate + ReturnDate + IsNonAnnual + Tuition + ToGrade + TotalSchoolEnrollment + FPP + SchoolSponsor + TotalPax + SingleGradeTripFlag, data = train) 

rf_model <- 
  rand_forest(mtry = tune(), min_n = tune(), trees = tune()) |> 
  set_engine("ranger") |> 
  set_mode("classification")

rf_workflow <- 
  workflow() |> 
  add_model(rf_model) |> 
  add_recipe(rf_recipe)

rf_res <- 
  rf_workflow |> 
  tune_grid(cv_set,
            grid = 25,
            control = control_grid(),
            metrics = metric_set(accuracy)) 

rf_res |>
  show_best(metric = "accuracy")
```

### Use the best parameters from the Random Forest model tuning to train the model

```{r p17}
rf_model <- 
  rand_forest(mtry = 3, min_n = 32, trees = 1093 ) |> 
  set_engine("ranger", importance = 'impurity') |> 
  set_mode("classification")

fit_workflow <- 
  workflow() |> 
  add_model(rf_model) |> 
  add_recipe(rf_recipe) |> 
  fit(data = train)


pred_class <- predict(fit_workflow,
                      new_data = test,
                      type = "class")

pred_probability <- predict(fit_workflow,
                      new_data = test,
                      type = "prob")

rf_results <- test |>
  bind_cols(pred_class, pred_probability)


custom_metrics(rf_results,
               truth = RetainedLabel,
               estimate = .pred_class)
```
### ROC Area Under the Curve

```{r p18}
roc_auc(rf_results,
        truth = RetainedLabel,
        .pred_Retained,
        event_level = "second")
```
### ROC Curve Visual

```{r p19}
rf_results |>
  roc_curve(truth = RetainedLabel, .pred_Retained, event_level = "second") |>
  autoplot()
```
### Variable Importance Plot

```{r p20}
fit_workflow |>
  extract_fit_parsnip() |> 
  vip(num_features = 20)
```

```{r p21}
model_data <- read_csv("scholastic_travel_predictions.csv", show_col_types = FALSE)
```


```{r p22}
pred_class <- predict(fit_workflow,
                      new_data = model_data,
                      type = "class")

model_data <- model_data |>
  mutate(RetainedLabel = pred_class$.pred_class)

write_csv(model_data, "model_data.csv")
```